---
title: "In-class_Ex4: SIMs"
author: "Widya Tantiya Yutika"
date: "09 December 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Overview

In this in-class exercise, you will gain hands-on experience on the following tasks:

-   performing geocoding using data downloaded from data.gov.sg

-   calibrating Geographically Weighted Poisson Regression

## Getting Started

```{r}
pacman::p_load(tidyverse, sf, httr, tmap)
```

Notes: httr allows us to work with html (to communicate with web server.

## Geocoding using SLA API

Address geocoding, or simply geocoding, is the process of taking a aspatial description of a location, such as an address or postcode, and returning geographic coordinates, frequently latitude/longitude pair, to identify a location on the Earth's surface.

Singapore Land Authority (SLA) supports online geocoding service called [OneMap API](https://www.onemap.gov.sg/apidocs/).

```{r}
#| eval: false
url <- "https://www.onemap.gov.sg/api/common/elastic/search"

csv <- read_csv("data/aspatial/Generalinformationofschools.csv")
postcodes <- csv$'postal_code'

found <- data.frame()
not_found <- data.frame()

for (postcode in postcodes){
  query <- list('searchVal'=postcode,'returnGeom'='Y','getAddrDetails'='Y','pageNum'='1')
  res<-GET(url, query=query)
  
  if((content(res)$found)!=0){
    found <-rbind(found, data.frame(content(res))[4:13])
  } else{
    not_found = data.frame(postcode)
  }
}
```

Notes:

#\| eval: false -\> run 1 time only

#\| echo: false -\> the code wont be shown on html page

#\| message: false -\> the message will not come up

```{r}
#| eval: false
merged = merge(csv, found, by.x='postal_code', by.y='results.POSTAL', all=TRUE)
write.csv (merged, file='data/aspatial/schools.csv')
write.csv (not_found, file ='data/aspatial/not_found.csv')
```

## Converting an aspatial data into a simple feature tibble data.frame

### Importing and tidying schools data

```{r}
schools <- read_csv("data/aspatial/schools.csv")%>%
  rename(latitude='results.LATITUDE', longitude='results.LONGITUDE')%>%
  select(postal_code, school_name, latitude,longitude)


```

### Converting an aspatial data into sf tibble data.frame

```{r}
schools_sf <- st_as_sf(schools, 
                       coords=c('longitude', 'latitude'),
                       crs=4326) %>%
  st_transform(crs=3414)
```

### Plotting a point simple feature layer

```{r}
tmap_mode("view")
tm_shape(schools_sf)+
  tm_dots()+
tm_view(set.zoom.limits =c(11,14))
tmap_mode("plot")
```

## Preparing

```{r}
mpsz <- st_read(dsn="data/geospatial", layer="MPSZ-2019")%>%
  st_transform(crs=3414)
```

```{r}
mpsz$'SCHOOL_COUNT' <- lengths(st_intersects(mpsz, schools_sf))
```

```{r}
summary(mpsz$SCHOOL_COUNT)
```

```{r}
business_sf <- st_read(dsn="data/geospatial", layer="Business")
```

```{r}
tmap_options(check.and.fix=TRUE)
tm_shape(mpsz)+
  tm_polygons()+
tm_shape(business_sf)+
  tm_dots()

```

```{r}
mpsz$'BUSINESS_COUNT' <- lengths(st_intersects(mpsz, business_sf))
summary(mpsz$BUSINESS_COUNT)
```

```{r}
#|eeval: false
#flow_data <- flow_data %>%
#  left_join(mpsz_tidy,
#            by =c("DESTIN_SZ"="SUBZONE_C"))
```

Notes: calibrate journey to home : then change to ORIGIN_SZ

## Checking for variables with zero values

## Model Calibration

```{r}
pacman::p_load(tmap, sf, performance, ggpubr, tidyverse)
```

## The Data

```{r}
flow_data<-read_rds("data/rds/flow_data_tidy.rds")
```

```{r}
glimpse(flow_data)
```

Notes: business small scale and middle scale industry; retail mainly for shopping purposes.

Notes: change to morning_peak to TRIPS and dist to DIST (because there is a dist function in R)

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ==flow_data$DESTIN_SZ,
  0, flow_data$MORNING_PEAK)
flow_data$offset <-ifelse(
  flow_data$ORIGIN_SZ==flow_data$DESTIN_SZ,
  0.000001,1
)
```

```{r}
inter_zonal_flow <- flow_data %>%
  filter(FlowNoIntra >0)
```

```{r}
inter_zonal_flow <- inter_zonal_flow %>%
  rename(TRIPS= MORNING_PEAK,
         DIST=dist)
```

### Origin(Production) constrained SIM

```{r}
orcSIM_Poisson <- glm(formula = TRIPS ~ 
                        ORIGIN_SZ +
                        log(SCHOOL_COUNT)+
                        log(RETAIL_COUNT)+
                        log(DIST) -1,
                      family = poisson (link='log'),
                      data = inter_zonal_flow,
                      na.action = na.exclude)
summary(orcSIM_Poisson)
```

Notes1: The purpose of -1 on log(DIST) is to remove the intercept.

Notes2: need to look at log(SCHOOL_COUNT) , log(RETAIL_COUNT), log(DIST), p-value

-   log(SCHOOL_COUNT) , log(RETAIL_COUNT) -\> attractiveness, need to be positive but if crime rate need to be negative (people dont want to stay near area with high crime rate)

```{=html}
<!-- -->
```
-   log(DIST) -\> need to be negative(inverse distance), meaning that the closer the distance, people are more willing to go

-   p-value \< 0.0.5 -\> if not \<0.05 need to recalibrate the model, remove the variables

### Goodness-of-Fit

```{r}
CalcRSquared <- function (observed, estimated){
  r<- cor(observed, estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(orcSIM_Poisson$data$TRIPS, orcSIM_Poisson$fitted.values)
```

```{r}
performance_rmse(orcSIM_Poisson, normalized= FALSE)
```

### Doubly Constrained

```{r}
dbcSIM_Poisson <- glm(formula = TRIPS ~
                        ORIGIN_SZ+
                        DESTIN_SZ+
                        log(DIST),
                      family = poisson(link='log'),
                      data= inter_zonal_flow,
                      na.action= na.exclude
                      )
dbcSIM_Poisson
```

```{r}
model_list <- list(originConstrained=orcSIM_Poisson, doublyConstrained=dbcSIM_Poisson)
```

```{r}
compare_performance(model_list, metrics="RMSE")
```

Notes: smaller RMSE -\> better model

Notes: if there is an outlier, try to remove and compare with current model.
