---
title: "In-class Exercise 1: My First Date with Geospatial Data Analytics"
author: "Widya Tantiya Yutika"
date: "18 November 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Overview

In this in-class exercise, I learn how to import and wrangling geospatial data using appropriate R packages.

## Getting Started

The code chunk below load the following packages:

-   tmap: for thematic mapping

-   sf: for geospatial data handling

-   tidyverse: non-spatial data handling

```{r}
pacman::p_load(tmap, sf, tidyverse)
```

## Preparing the Flow Data

## Importing the OD Data

Firstly, we will import the *Passenger Volume by Origin Destination Bus Stops* data set downloaded from LTA DataMall by using read_csv() of **readr** package.

```{r}
# eval:false
odbus <- read_csv("data/aspatial/origin_destination_bus_202308.csv")
```

### Extracting the study data

```{r}
#eval: false
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE)
```

```{r}
# eval: false
origtrip_7_9 <- odbus %>%
  filter(DAY_TYPE=="WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 &
           TIME_PER_HOUR <=9) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS=sum(TOTAL_TRIPS))
```

Two geospatial data will be used in this exercise, they are:

```{r}
busstop <- st_read(dsn="data/geospatial", layer="BusStop") %>%
  st_transform(crs=3414)
```

```{r}
mpsz <- st_read(dsn="data/geospatial", layer="MPSZ-2019") %>%
  st_transform(crs=3414)
mpsz
```
