---
title: "In-class_Ex2_ver1"
author: "Widya Tantiya Yutika"
date: "25 November 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Getting Started

### Installing and Loading the R Packages

FiveR packages will be used for this in-class exercise, they are: sf, sfdep, tmap, tidyverse, and knitr.

```{r} pacman::p_load(sf, spdep, tmap, tidyverse, knitr)}
```

## The Data

For the purpose of this in-class exercise, the Hunan data sets will be used. There are two data sets in this use case, they are:

-   Hunan, a geospatial data set in ESRI shapefile format, and

-   Hunan_2012, an attribute data set in csv format.

## Getting Data in R Environment

### **Importing geospatial data**

```{r} hunan <- st_read(dsn = "data/geospatial", layer = "Hunan")}
```

### **Importing attribute table**

```{r} hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")}
```

### Combining both data frame by using left join

```{r} hunan_GDPPC <- left_join(hunan,hunan2012)%>%   select(1:4, 7, 15)}
```

In order to retain the geospatial properties, the left data frame must be the sf data.frame(i.e. hunan)

### Plotting a choropleth map

```{r}
```

## Deriving Continuity Spatial Weights

## Deriving Continuity Spatial Weights: Queen's Method

In the code below, queen method is used to derive the contiguity weights.

```{r} wm_q <- hunan_GDPPC %>%   mutate(nb = st_contiguity(geometry),          wt = st_weights(nb,                          style='W'),          .before=1)}
```

Notes: ,before1 -\> put nb and wt at the front of the tibble dataset
