---
title: "In-class_Ex3:Calibrating Spatial Interaction Models with R"
author: "Widya Tantiya Yutika"
date: "02 December 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Overview

Spatial Interaction Models (SIMs) are mathematical models for estimating flows between spatial entities developed by Alan Wilson in the late 1960s and early 1970, with considerable uptake and refinement for transport modelling since then Boyce and Williams (2015).

There are four main types of traditional SIMs (Wilson 1971):

-   Unconstrained

-   Production-constrained

-   Attraction-constrained

-   Doubly-constrained

Ordinary least square (OLS), log-normal, Poisson and negative binomial (NB) regression methods have been used extensively to calibrate OD flow models by processing flow data as different types of dependent variables. In this chapter, you will gain hands-on experiences on using appropriate R packages to calibrate SIM by using there four regression methods.

## **The Case Study and Data**

In this exercise, we are going to calibrate SIM to determine factors affecting the public bus passenger flows during the morning peak in Singapore.

## **Getting Started**

For the purpose of this exercise, four r packages will be used. They are:

-   sf for importing, integrating, processing and transforming geospatial data.

-   tidyverse for importing, integrating, wrangling and visualising data.

-   tmap for creating thematic maps.

-   DT: for dynamic table

-   sp: for spatial data handling

-   performance: for creating performance objects

-   reshape2: for reshaping data

-   ggpubr: for creating multiple plot into 1

```{r}
pacman::p_load(tmap, sf, sp, DT,
               performance, reshape2,
               ggpubr, tidyverse)
```

## **The Data**

This exercise is a continuation of **Chapter 15: Processing and Visualising Flow Data** and the following data will be used:

-   *od_data.rds*, weekday morning peak passenger flows at planning subzone level.

-   *mpsz.rds*, URA Master Plan 2019 Planning Subzone boundary in simple feature tibble data frame format.

Beside these two data sets, an additional attribute data file called pop.csv will be provided.

## **Computing Distance Matrix**

In spatial interaction, a distance matrix is a table that shows the distance between pairs of locations. For example, in the table below we can see an Euclidean distance of 3926.0025 between MESZ01 and RVSZ05, of 3939.1079 between MESZ01 and SRSZ01, and so on. By definition, an location\'s distance from itself, which is shown in the main diagonal of the table, is 0.

In this section, you will learn how to compute a distance matrix by using URA Master Plan 2019 Planning Subzone boundary in which you saved as an rds file called *mpsz*.

First, let us import *mpsz.rds* into R environemnt by using the code chunk below.

```{r}
mpsz <- read_rds("data/rds/mpsz.rds")
mpsz
```

Notice that it is a sf tibble dataframe object class.

### **Converting from sf data.table to SpatialPolygonsDataFrame**

There are at least two ways to compute the required distance matrix. One is based on sf and the other is based on sp. Past experience shown that computing distance matrix by using sf function took relatively longer time that sp method especially the data set is large. In view of this, sp method is used in the code chunks below.

First [`as.Spatial()`](https://r-spatial.github.io/sf/reference/coerce-methods.html) will be used to convert *mpsz* from sf tibble data frame to SpatialPolygonsDataFrame of sp object as shown in the code chunk below.

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

```{r}
#mpsz_sp_selected <- mpsz_sp %>%
#  select(mpsz@data$SUBZONE)
```

### **Computing the distance matrix**

Next, [`spDists()`](https://www.rdocumentation.org/packages/sp/versions/2.1-1/topics/spDistsN1) of sp package will be used to compute the Euclidean distance between the centroids of the planning subzones.

::: callout-tip
### Q&A

Do you know why the distance is calculated between two centroids of a pair of spatial polygons?
:::

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
head(dist, n=c(10, 10))
```

Notice that the output *dist* is a matrix object class of R. Also notice that the column heanders and row headers are not labeled with the planning subzone codes.

### **Labelling column and row heanders of a distance matrix**

First, we will create a list sorted according to the the distance matrix by planning sub-zone code.

```{r}
sz_names <- mpsz$SUBZONE_C
```
